Packaging research artefacts with RO-Crate

### 2022-01-02 latex/ro-crate.tex
Formatting in appendix

An important RO-Crate principle is that of being \textbf
{self-described} Therefore the serialisation of the RO-Crate into a
file should also describe itself in a Metadata File Descriptor,\footnote{\url
{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}} indicating it is $\textit{about}$ (describing) the RO-Crate root data entity,
and that it $\textit{conformsTo}$ a particular version of the RO-Crate specification:
%
\begin{eqnarray*}
\textit{about}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
{<http://schema.org/about>}, o)
\\
\textit{conformsTo}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
{<http://purl.org/dc/terms/conformsTo>}, o)
\\
\textit{MetadataFile}(m) & \Rightarrow& \textit{CreativeWork}(m) \land
\textit{about}(m,R) {\wedge } \textit{ROCrate}(R) \land
\\
& &{} \textit{conformsTo}(m, \texttt{<https://w3id.org/ro/crate/1.1>})
\end{eqnarray*}

@@

In the conceptual model the \emph{RO-Crate Metadata File} can be seen
as the top-level node that describes the \emph{RO-Crate Root}, however
in the formal model (and the JSON-LD format) the metadata file
descriptor is an additional contextual entity that is not affecting the
depth-limit of the RO-Crate.

@@

Combining the above predicates and Schema.org mapping with rudimentary
JSON templates, these forward-chaining production rules can output
JSON-LD according to the RO-Crate 1.1 specification:\footnote{\textbf{Limitations:}
 Contextual entities not related from the
RO-Crate (e.g. using inverse relations to a data entity) would not be
covered by the single direction $Mentions(R, s)$ production rule; see
\href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub
issue ResearchObject/ro-crate\#122}. The $\textit{datePublished}(e, d)$ rule do
not include syntax checks for the ISO 8601 datetime format. Compared
with RO-Crate examples, this generated JSON-LD does not use a
$\textit{@context}$ as the IRIs are produced unshortened, a post-step could do
JSON-LD Flattening with a versioned RO-Crate context. The \texttt
{@type} expansion is included for clarity, even though this is also
implied by the $\textit{type}(e, t)$ expansion to $\textit{Relation}(e, \texttt{xsd:type})$.}
%
\begin{eqnarray*}
\textit{Mentions}(R, s) \land \textit{Relation}(s,p,o) & \Rightarrow&
\textit{Mentions}(R, o)
\\
\textit{IRI}(i) & \Rightarrow& \texttt{"} i \texttt{"}
\\
\textit{Decimal}(d) & \Rightarrow& d
\\
\textit{String}(s) & \Rightarrow& \texttt{"} s \texttt{"}
\\
\forall e \forall t \ \textit{type}(e, t) & \Rightarrow& \texttt{\{"@id":} e
\texttt{,}
\\
&& \ \ \texttt{"@type":} t
\\
&& \texttt{\}}
\\
\forall s \forall p \forall o \ \textit{Relation}(s,p,o) & \Rightarrow& \texttt{
\{"@id":} s \texttt{,}
\\
&& \ \ p \texttt{: \{ "@id":} o \texttt{\}}
\\
&& \texttt{\}}
\\
\forall s \forall p \forall v \ \textit{Attribute}(s,p,v) & \Rightarrow&
\texttt{\{"@id":} s \texttt{,}
\\
&& \ \ p \texttt{:} v
\\
&& \texttt{\}}
\\
\forall r \forall c\ \textit{ROCrate}(r) & \Rightarrow& \texttt{\{
"@graph": [}
\\
&& \ \ \ \ \textit{Mentions}(r, c)*
\\
&& \ \ \ \texttt{]}
\\
&& \texttt{\}}
\\
R & \equiv& \texttt{<./>}
\\
R & \Rightarrow& \textit{MetadataFile}(\texttt{<ro-crate-metadata.json>})
\end{eqnarray*}

>>>>

An important RO-Crate principle is that of being \textbf
{self-described}. Therefore the serialisation of the RO-Crate into a
file should also describe itself in a Metadata File Descriptor,\footnote{\url
{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}} indicating it is $\textit{about}$ (describing) the RO-Crate root data entity,
and that it $\textit{conformsTo}$ a particular version of the RO-Crate specification:
%
\begin{eqnarray*}
\textit{about}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
{<http://schema.org/about>}, o)
\\
\textit{conformsTo}(s,o) & \Rightarrow& \textit{Relation}(s, \texttt
{<http://purl.org/dc/terms/conformsTo>}, o)
\\
\textit{MetadataFile}(m) & \Rightarrow& \textit{CreativeWork}(m) \land
\textit{about}(m,R) \land \textit{ROCrate}(R) \land
\\
& &{} \textit{conformsTo}(m, \texttt{<https://w3id.org/ro/crate/1.1>})
\end{eqnarray*}

@@

In the conceptual model, the \emph{RO-Crate Metadata File} can be seen
as the top-level node that describes the \emph{RO-Crate Root}, however
in the formal model (and the JSON-LD format) the metadata file
descriptor is an additional contextual entity that is not affecting the
depth-limit of the RO-Crate.

@@

Combining the above predicates and Schema.org mapping with rudimentary
JSON templates, these forward-chaining production rules can output
JSON-LD according to the RO-Crate 1.1 specification:\footnote{\textbf{Limitations:}
 Contextual entities not related from the
RO-Crate (e.g. using inverse relations to a data entity) would not be
covered by the single direction $Mentions(R, s)$ production rule; see
\href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub
issue ResearchObject/ro-crate\#122}. The $\textit{datePublished}(e, d)$ rule do
not include syntax checks for the ISO 8601 datetime format. Compared
with RO-Crate examples, this generated JSON-LD does not use a
$\textit{@context}$ as the IRIs are produced unshortened, a post-step could do
JSON-LD Flattening with a versioned RO-Crate context. The \texttt
{@type} expansion is included for clarity, even though this is also
implied by the $\textit{type}(e, t)$ expansion to $\textit{Relation}(e, \texttt{xsd:type})$.}
%
\begin{eqnarray*}
\textit{Mentions}(R, s) \land \textit{Relation}(s,p,o) & \Rightarrow&
\textit{Mentions}(R, o)
\\
\textit{IRI}(i) & \Rightarrow& \texttt{"} i \texttt{"}
\\
\textit{Decimal}(d) & \Rightarrow& d
\\
\textit{String}(s) & \Rightarrow& \texttt{"} s \texttt{"}
\\
\forall e \forall t \ \textit{type}(e, t) & \Rightarrow& \texttt{\{"@id":} e
\texttt{,}
\\
&& \ \ \texttt{"@type":} t
\\
&& \texttt{\}}
\\
\forall s \forall p \forall o \ \textit{Relation}(s,p,o) & \Rightarrow& 
\texttt{\{"@id":} s \texttt{,}
\\
&& \ \ p \texttt{: \{"@id":} o \texttt{\}}
\\
&& \texttt{\}}
\\
\forall s \forall p \forall v \ \textit{Attribute}(s,p,v) & \Rightarrow&
\texttt{\{"@id":} s \texttt{,}
\\
&& \ \ p \texttt{:} v
\\
&& \texttt{\}}
\\
\forall r \forall c\ \textit{ROCrate}(r) & \Rightarrow& \texttt{\{
"@graph": [}
\\
&& \ \ \ \ \ \ \textit{Mentions}(r, c)*
\\
&& \ \ \ \texttt{]}
\\
&& \texttt{\}}
\\
R & \equiv& \texttt{<./>}
\\
R & \Rightarrow& \textit{MetadataFile}(\texttt{<ro-crate-metadata.json>})
\end{eqnarray*}
